### 2019-09-16 appendix.Rmd
Appendix added.

---
title: "Appendix"
type: Short Technical Note
author:
  - name: Ursula Laa
    affil: a, b
    email: ursula.laa@monash.edu
  - name: Dianne Cook
    affil: b
    email: dicook@monash.edu
  - name: German Valencia
    affil: a
    email: german.valencia@monash.edu
affiliation:
  - num: a
    address: |
      School of Physics and Astronomy, Monash University
  - num: b
    address: |
      Department of Econometrics and Business Statistics, Monash University
bibliography: biblio.bib
geometry: margin=2.5cm
abstract: |
  Taking projections of high-dimensional data is a common analytical and visualisation technique in statistics for working with high-dimensional problems. Sectioning, or slicing, through high dimensions is less common, but can be useful for visualising data with concavities, or non-linear structure. It is associated with conditional distributions in statistics, and also linked brushing between plots in interactive data visualisation. This short technical note describes a simple approach for  slicing in the orthogonal
  space of projections obtained when running a tour, thus presenting the viewer
  with an interpolated sequence of sliced projections. The method has been
  implemented in R as an extension to the tourr package, and can be used to
  explore for concave and non-linear structures in multivariate distributions.
keywords: |
  data visualisation; grand tour; sectioning
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
  \usepackage{setspace}
output: rticles::tf_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

\doublespacing
# Slice volume derivation

We calculate the relative volume of a slice through a $p$ dimensional hypersphere. The volume of a full hypersphere with radius $r$ is

\begin{equation}
V(r, p) = \frac{\pi^{p/2} r^p}{\frac{p}{2} \Gamma(p/2)}.
\end{equation}

It is parametrised by its radius $r$ and its variation with the radius is given as

\begin{equation}
\frac{dV(r, p)}{dr} = 2 \frac{\pi^{p/2} r^{p-1}}{\Gamma(p/2)}.
\end{equation}

Our slice definition is spherical in the $p-2$ dimensional orthogonal space, while capturing the full volume (i.e. area) inside the plane. To calculate the slice volume we therefore integrate $\frac{dV(r, p-2)}{dr}$ multiplied with the area in the plane, from the origin to the slice thickness $h$. Using the parametrisation with the radius $r$, the area in the plane is a circle with radius $\sqrt{R^2 - r^2}$. The slice volume is thus calculated as

\begin{equation}
V_{slice}(R, h, p) =
\int_0^h \frac{dV(r, p)}{dr} V(\sqrt{R^2 - r^2}, 2) dr =
\frac{\pi^{p/2}}{\Gamma(p/2)} \frac{h^{p-2}}{p} [pR^2 - (p-2)h^2].
\end{equation}

The relative volume of the slice is then simply given as

\begin{equation}
V_{rel}(R, h, p) = \frac{V_{slice}(R, h, p)}{V(R, p)} = \frac{h^{p-2}}{2R^p} [pR^2 - (p-2)h^2].
\end{equation}