### 2019-09-18 coverletter.Rmd
Coverletter added.

---
from: Ursula Laa
fromqualifications: PhD
fromtitle: Research fellow
fromemail: ursula.laa@monash.edu
fromWWW: github.com/uschiLaa
fromphone: +61 4 78 001 984
fromsignature: signature.pdf
toname: Editor
toaddress: Journal of Computational and Graphical Statistics
opening: Dear Dr McCormick,
closing: Sincerely
spacing: 1.4
output: MonashEBSTemplates::letter
---

Please consider our manuscript titled "A slice tour for finding hollowness in high-dimensional data" for publication as a short technical note in the Journal of Computational and Graphical Statistics.

This note describes a new method for the visualisation of high-dimensional data by defining slices in the orthogonal space of 2D projections. It can be combined with a tour for a dynamic display that shows slices on a path of interpolated projections. The method has been implemented in R as a new display option in the tourr package. It is particularly useful to explore for concave and non-linear structures in multivariate distributions. Several examples are included in the note and illustrate the potential for different types of applications.

We believe that it is an ideal fit for a short technical note in the Journal of Computational and Graphical Statistics. It describes a new graphical method that will be useful to people working in multivariate statistics.

Thank you for the consideration of this manuscript.