### 2019-09-23 appendix.Rmd
Appendix format changed.

---
title: "Appendix"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

# Slice volume derivation

We calculate the relative volume of a slice through a $p$ dimensional hypersphere of radius $R$. In general the volume of a full hypersphere with radius $r$ in $q$ dimensions is given as
\begin{equation}
V(r, q) = \frac{\pi^{q/2} r^q}{\frac{q}{2} \Gamma(q/2)},
\end{equation}
and its variation with the radius is
\begin{equation}
\frac{dV(r, q)}{dr} = 2 \frac{\pi^{q/2} r^{q-1}}{\Gamma(q/2)}.
\end{equation}

To calculate the slice volume we note that it is spherical in the $p-2$ dimensional orthogonal space on the projection plane, while capturing the full volume (i.e. area) inside the plane. To calculate the slice volume we therefore integrate the product of $\frac{dV(r, p-2)}{dr}$ and the area in the plane parametrised by $r$, from the origin to the full slice thickness $h$. The area in the plane is a circle with radius $\sqrt{R^2 - r^2}$. The slice volume is thus calculated as
\begin{equation}
V_{slice}(R, h, p) =
\int_0^h \frac{dV(r, p-2)}{dr} V(\sqrt{R^2 - r^2}, 2) dr =
\frac{\pi^{p/2}}{\Gamma(p/2)} \frac{h^{p-2}}{p} [pR^2 - (p-2)h^2].
\end{equation}
The relative volume of the slice is then simply given as
\begin{equation}
V_{rel}(R, h, p) = \frac{V_{slice}(R, h, p)}{V(R, p)} = \frac{h^{p-2}}{2R^p} [pR^2 - (p-2)h^2].
\end{equation}